Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2024-03-13 2_SSL_NE/article/article_supplementary.tex
updated figures and captions

% custom section designations
\makeatletter
\renewcommand{\@seccntformat}[1]{Supplementary Note \csname the#1\endcsname: } % https://tex.stackexchange.com/a/693618
\makeatother
% custom equation designations
\renewcommand{\figurename}{Supplementary Figure}
\crefname{figure}{Supplementary Figure}{Supplementary Figures}
\renewcommand{\tablename}{Supplementary Table}
\crefname{table}{Supplementary Table}{Supplementary Tables}

@@

\clearpage
\section{Previous Literature on Techological Progress in LEDs}
\label{sec:prev_lit}

@@

\section{Metrics for Technological Progress in LEDs}

This Supplementary Note provides additional background and details for the physical, economic and consumer experience metrics that were chosen to quantify multi-dimensional progress in light-emitting diode technology.

\subsection{Overview}

@@

\subsection{Device Performance Metrics}
\label{sec:device_performance_metrics}

@@

\subsubsection{Forward Voltage Efficiency $\eta_{V_f}$}

@@

\subsubsection{Light Extraction Efficiency $\eta_{LE}$}

@@

\subsubsection{Internal Quantum Efficiency $\eta_{IQ}$}

@@

\subsubsection{Droop $\eta_{droop}$}

@@

\subsubsection{Light Conversion Efficiency $\eta_{C}$}

@@

\subsubsection{Spectral Efficiency $\eta_{S}$}

@@

\subsubsection{Luminous Efficacy of Radiation}
\label{subsec:ler}

@@

\subsubsection{Luminous Efficacy of Source}
\label{subsec:les}

@@

\subsection{Consumer Experience Metrics}

@@

\subsubsection{Color Rendering Index (CRI)}

@@

\subsubsection{Colour Temperature}

@@

\subsection{Cost Metrics}

@@

\clearpage
\section{Methodological Details}
\label{sec:method_details}

This Supplementary Note provides additional details for the main methods and data sources used in our mixed-methods research approach described in the Methods section in the main text.

\subsection{Background of Interviewed Experts}

\begin{table}[H]
\small
    \centering
    \caption{Anonymized list of interviewed LED experts.}
    \vspace{5mm}
    \begin{tabular}{|l|l|l|l|l|}
    \hline
        \textit{\#} & \textit{Sector} & \textit{Role} & \textit{Country} & \textit{Area of Expertise} \\ \hline
        1 & Academia & Senior researcher & UK & Epitaxy \\ \hline
        2 & Industry & Consultant, former senior researcher & USA & Device architecture \\ \hline
        3 & Industry & Consultant, former head of R\&D & Germany & Epitaxy \\ \hline
        4 & Academia & Professor & Austria & Phosphors \\ \hline
        5 & Industry & Consultant, former head of R\&D & USA & Device architecture \\ \hline
        6 & Consulting & Consultant, former senior technical advisor & USA & Device architecture \\ \hline
        7 & Academia & Professor & Germany & Phosphors \\ \hline
        8 & Government & R\&D manager & USA & Device architecture \\ \hline
        9 & Consulting & Consultant & USA & Device applications \\ \hline
        10 & Academia & Professor & France & Device physics \\ \hline
        11 & Industry & Senior scientist, former head of R\&D & USA & Device architecture \\ \hline
        12 & Industry & Principal scientist & Germany & Phosphors \\ \hline
        13 & Industry & Former head of R\&D & USA & Phosphors \\ \hline
    \end{tabular}
    \label{tab:interviews}
\end{table}

\subsection{Manufacturing Cost Model Details}
\label{sec:costmodel_details}

\subsubsection{Structure of the Model}

@@

The LED cost model developed for this publication is a microeconomic manufacturing cost model. Within the timeframe and scope laid out in the main text of this publication, it returns the total manufacturing cost of GaN-based phosphor-converted warm white light-emitting diode packages. The model considers main economic factors associated with operating and maintaining manufacturing equipment. It does not consider costs associated with research and development or those associated with the construction of manufacturing facilities. It considers market trends through their effect on manufacturing parameters.

@@

\subsubsection{Manufacturing Process Steps by Chip Architecture}
\label{subsec:manufacturing_steps}

@@

    \vspace{5mm}

@@

\newpage
\section{Cost Model Sensitivity Analysis}

\begin{figure}[ht!]
	\centering
    \includegraphics[width=\textwidth]{./figures/costmodel_sensitivity.pdf}
	\caption{Sensitivity analysis for selected cost model parameters. Panel a represents the tested range of variation of parameters indicated on horizontal axis along with their respective measurement units, as represented in Supplementary Table 3. Panel b shows the resulting range of variation in the manufacturing cost. Ranges are represented for three years considered in the cost model, where corresponding data is available, by whiskers of the following colour: 2003 – blue, 2012 – green, 2020 – red. Note the overall trend of decreasing sensitivity to cost model parameters as the number of die per wafer increases over time. Abbreviations: resp. - respective; Yield Sav. from Insp. - yield savings from inspection.}
	\label{fig:sensitivity}
\end{figure}

@@

\begin{table}[]
\small
    \caption{List of LED manufacturing cost model parameters considered in sensitivity analysis}
    \begin{NiceTabularX}{1.1\textwidth}{ |X|X|l|l|l|l|l|l|X|}
        \hline
            \textit{Parameter} & \textit{Unit} & \textit{2003} & $\pm [\%]$ & \textit{2012} & $\pm [\%]$ & \textit{2020} & $\pm [\%]$ & Source \\
        \hline
            Cleanroom Cost & \text{USD}/m$^2$ & 3000 & +16,-16 & 3000 & +16,-16 & 3000 & +16,-16 & \cite{mddi1997cleanroom}\cite{ledcomv2} \newline \cite{bakshi2009euv}\cite{gajera2006process} \\
        \hline
            Labor Rate & \$/h & 30 & +50,-50 & 30 & +50,-50 & 30 & +50,-50 & I \\
        \hline
            Equip. Discount & \% of \text{USD} & 0\% & +25,-25 & 0\% & +25,-25 & 0\% & +25,-25 & I \newline \cite{Appleyard_2001} \\
        \hline
            Overall Yield & \% & 31\% & +25,-25 & 57\% & +25,-15 & 74\% & +25,-25 & I, \cite{lumi2012yield}\cite{ledsmag2012} \newline \cite{systemplus2015reverse}\cite{ledcomv2} \\
        \hline
            Inspec. Yield Savings & \%/inspec. & 0.5\% & +80,-80 & 0.5\% & +80,-80 & 0.5\% & +80,-80 & \cite{mckinseyyield} \\
        \hline
            Overall Throughput & UPH or h$^{-1}$ & varies & +50,-50 & varies & +50,-50 & varies & +50,-50 & Equipment Datasheets \\
        \hline
            Wafer Diameter & mm & 100 & +0,-50 & 150 & +33.3,-33.3 & 200 & +0,-25 & I and \cref{fig:wafers} \\
        \hline
            Edge Exclusion & mm & 7 & +0,-50 & 5 & +40,-0 & 5 & +40,-20 & \cite{ledsmagexclusion}\cite{rubiconexclusion} \newline \cite{xiamenexclusion}\cite{american2007annual} \\
        \hline
            Cutting Width & $\mu$m & 100 & +50,-25 & 75 & +33.3,-20 & 20 & +300,-50 & \cite{masaki2000division}\cite{ils2005width} \newline \cite{photonics2010width}\cite{discowidth} \\
        \hline
            Metal Prices & \text{USD}/kg & varies & +50,-0 & varies & +50,-0 & varies & +50,-0 & Datasheets \\
        \hline
            Electricity Price & \text{USD}/kWh & 0.07 & +50,-50 & 0.08 & +50,-50 & 0.06 & +50,-50 & \cite{eia2000electric}\cite{eia2019electric} \\
        \hline
            Saph. Subst. Price & \text{USD} & 40 & +12.5,-12.5 & 10 & +100,-20 & 3 & +66.6,-33.3 & \cref{fig:wafers} \\
        \hline
            Phosphor Prices & \text{USD}/g & varies & +50,-50 & varies & +0,-0 & varies & +0,-0 & I, \cite{yole_phosphor_2012}\cite{yole2017phosphor} \\
        \hline
        \end{NiceTabularX}
    \caption*{The results of the sensitivity analysis for the parameters in this table are presented in \cref{fig:sensitivity}. Note: Baseline values are provided only for \textit{global} model parameters in the columns \textit{2003, 2012, 2020}. For \textit{non-global} model parameters (those which vary by equipment) only the variation, in percent of the individual baseline values is provided. The corresponding units for the values in columns \textit{2003}-\textit{2020} are indicated in the column \textit{Units}. The variation employed during the sensitivity analysis is provided in the columns $\pm [\%]$. Abbreviations: UPH - units per hour; Equip. Discount - equipment discount (sales rebate for large purchases); Inspec. Yield Savings - yield savings from inspection (early detection and alleviation of issues in the manufacturing workflow); Saph. Subst. - Sapphire Substrate. ’I’ in the Source column indicates expert interviews as the source of information.}
    \label{tab:sensitivity}
\end{table}

\clearpage
\section{Cost Model Comparison with Industry Data and DOE Projections}

We conduct a comparison of the outcomes of our LED manufacturing cost model regarding the evolution of the cost structure of low-to-mid-power classic chip GaN phosphor-converted white LED with previously reported U.S. DOE calculations and projections based on the LEDCOM model and industry data provided as part of industry round table discussions (see \cref{fig:costmodel_calibration}). We note some differences between the results of our model and the cost structure reported or projected by the DOE. For instance, the share of the epitaxy step is consistently larger in the DOE data. This can in part be explained by our model relying on state-of-the-art equipment at a virtual U.S.-based manufacturing location, while industry might not run low-power and mid-power chip production on these, more expensive, reactors. In addition, the manufacturing lines of the majority of manufacturers are located in Asia. We also note that the share of the sapphire substrate price in the DOE data is much larger than in our model, which can in part be explained by the overestimation of the actual price of sapphire wafers in earlier projections made for 2015-2020. Finally, the relative importance of the packaging part of the manufacturing process is very similar in our model and in the DOE results in 2012. However, in the DOE projections it significantly decreases by 2020, while in our model it retains and even increases its share of the total cost. This trend has been independently confirmed by researchers and industry reports on wafer-level packaging \cite{Lee2011WPL,Xie2013,ledsmag2017WLP}, showing better performance of our cost model compared to earlier DOE model projections.

@@

\clearpage
\section{List of Sources for Figures}
\label{sec:sources}

\subsection{Figure 1 (Historical Progress in Luminous Efficacy of Lighting Technologies)}

\begin{table}[h!]
    \caption{List of sources for Figure 1 organized by technology.}
    \begin{NiceTabularX}{\textwidth}{|l|X|}
    \hline
    \textit{Technology} & \textit{References} \\
    \hline
    LED & Own research, compare \cite{zenodo_weinold_led_history} \\
    \hline
    CFL (<1984) & \cite{Bouwknegt1982}\cite{Vrenken1983} \\
    \hline
    CFL (1984-2011) & \cite{eger2018origin} \\
    \hline
    CFL (>2011) & \cite{Guan2015} \\
    \hline
    Fire, Incandescent, HID & \cite{azevedo2009transition} augmented by own calculations based on \cite{benesch1905beleuchtungswesen} \\
    \hline
    Max. efficacy & \cite{Murphy2012} \\
    \hline
    \end{NiceTabularX}
    \caption*{Note: CFL - Compact Fluorescent Lamp. HID - High-Intensity Discharge.}
\end{table}

\subsection{Figure 2 (Historical Development of LED Lamp Prices)}

The sources for Figure 2 in the main article are grouped by the provider of data:

@@

\subsection{Figure 3 (Historical Evolution of LED Chip Architectures)}

@@

\subsection{\cref{fgr:subeff} (Historical Developments in LED Sub-Efficiencies)}

@@

\printbibliography

>>>>

% custom section designations
% https://tex.stackexchange.com/a/712523
\titleformat{\section}[display]{\large\bfseries\filright}{\bfseries Supplementary Note \thesection}{0pt}{}
\crefname{section}{Supplementary Note}{Supplementary Notes}
% custom equation designations
\renewcommand{\figurename}{Supplementary Figure}
\crefname{figure}{Supplementary Figure}{Supplementary Figures}
\renewcommand{\tablename}{Supplementary Table}
\crefname{table}{Supplementary Table}{Supplementary Tables}

@@

\clearpage

\section{Previous Literature on Techological Progress in LEDs}
\label{sec:prev_lit}

@@

\section{Metrics: Technological Progress in LEDs}

@@

\section{Metrics: Device Performance}
\label{sec:device_performance_metrics}

@@

\textbf{Forward Voltage Efficiency $\eta_{V_f}$}

@@

\textbf{Light Extraction Efficiency $\eta_{LE}$}

@@

\textbf{Internal Quantum Efficiency $\eta_{IQ}$}

@@

\textbf{Droop $\eta_{droop}$}

@@

\textbf{Light Conversion Efficiency $\eta_{C}$}

@@

\textbf{Spectral Efficiency $\eta_{S}$}

@@

\textbf{Luminous Efficacy of Radiation}
\label{subsec:ler}

@@

\textbf{Luminous Efficacy of Source}
\label{subsec:les}

@@

\clearpage
\section{Metrics: Consumer Experience}

@@

\textbf{Color Rendering Index (CRI)}

@@

\textbf{Colour Temperature}

@@

\textbf{Cost Metrics}

@@

\clearpage
\section{Methods: Background of Interviewed Experts}

\begin{table}[H]
\small
    \centering
    \caption{Anonymized list of interviewed LED experts.}
    \begin{tabular}{|l|l|l|l|l|}
    \hline
        \textit{\#} & \textit{Sector} & \textit{Role} & \textit{Country} & \textit{Area of Expertise} \\ \hline
        1 & Academia & Senior researcher & UK & Epitaxy \\ \hline
        2 & Industry & Consultant, former senior researcher & USA & Device architecture \\ \hline
        3 & Industry & Consultant, former head of R\&D & Germany & Epitaxy \\ \hline
        4 & Academia & Professor & Austria & Phosphors \\ \hline
        5 & Industry & Consultant, former head of R\&D & USA & Device architecture \\ \hline
        6 & Consulting & Consultant, former senior technical advisor & USA & Device architecture \\ \hline
        7 & Academia & Professor & Germany & Phosphors \\ \hline
        8 & Government & R\&D manager & USA & Device architecture \\ \hline
        9 & Consulting & Consultant & USA & Device applications \\ \hline
        10 & Academia & Professor & France & Device physics \\ \hline
        11 & Industry & Senior scientist, former head of R\&D & USA & Device architecture \\ \hline
        12 & Industry & Principal scientist & Germany & Phosphors \\ \hline
        13 & Industry & Former head of R\&D & USA & Phosphors \\ \hline
    \end{tabular}
    \label{tab:interviews}
\end{table}

\clearpage
\section{Methods: Cost Model Structure}
\label{sec:costmodel_structure}

@@

The LED cost model developed for this publication is a bottom-up microeconomic manufacturing cost model with process-step resolution. Within the timeframe and scope laid out in the main text of this publication, it returns the total manufacturing cost of GaN-based phosphor-converted warm white light-emitting diode packages. The model assumes production at a \textit{"virtual"} North American factory which employs the highest performing equipment which could be identified. Performance here includes throughput, associated yield and resource consumption. It assumes North American utility prices as well as labour rates, which have remained constant over the time period of interest. The equipment used to model the manufacturing processes has been chosen from the catalogues of current market leaders in semiconductor equipment manufacturing. This choice was made based on the specialization of the equipment, as some manufacturers offer product lines aimed specifically at light-emitting diode manufacturing. When more than one type of manufacturing equipment was available, the higher volume variant was chosen. The model does not consider costs associated with research and development or those associated with the construction of manufacturing facilities. It considers market trends through their effect on manufacturing parameters.

@@

To provide some context on the scope of our model, a selection of similar efforts is listed in \cref{tab:costmodels}:

\begin{table}[h!]
    \caption{\textbf{Recent manufacturing cost models for clean energy technologies.}}
    \begin{NiceTabularX}{\textwidth}{ |l|l|X|l|l|l|l|l|l|l| }
    \CodeBefore
       \rectanglecolor{blue!10}{13-1}{14-10}
    \Body
    \hline
        \multirow{2}{*}{\textit{Year}} & \multirow{2}{*}{\textit{Tech.}} & \multirow{2}{*}{\textit{Purpose}} & \multirow{2}{*}{\textit{Period}} & \multirow{2}{*}{\textit{Base}} &  \multicolumn{4}{c|}{\textit{Parameters}} & \multirow{2}{*}{\textit{Source}} \\
    \cline{6-9}
         & & & & & \textit{M} & \textit{E} & \textit{G} & \textit{S} & \\
    \hline
        2018 & Solar & Cost reductions & 1980-2012 & steps & Y & N & Y & Y* & \cite{kavlak2018evaluating} \\
    \hline
        2013 & Solar & Geographic diff. in Cost & 2012 & Red. & Y & Y & Y & N & \cite{goodrich2013assessing} \\
    \hline
        2006 & Solar & Cost reductions & 1975-2001 & Red. & Y & N & Y & N & \cite{Nemetsolar} \\
    \hline
        2011 & IC & Cost reductions & 2011 & Steps & N & N & Y & Y & \cite{Gopalakrishnan2011} \\
    \hline
        2017 & Li-Ion & Technology comparison & 2017 & Steps & (Y) & (Y) & Y & Y & \cite{Ciez2017} \\
    \hline
        2015 & Li-Ion & Industry & 2015 & Steps & N & N & Y & Y &  \cite{Petri2015} \\
    \hline
        2012 & Li-Ion & Manufacturers & 2012-now & Steps & (Y) & N & Y & Y & \cite{Nelson2012} \\
    \hline
        2020 & Wind & Impact of breakthroughs & 2005-2017 & Steps & Y & Y & Y & Y & \cite{Elia2020} \\
    \hline
        2010 & Wind & Future cost savings & 2010 & Steps & N & N & Y & Y & \cite{Schubel2010} \\
    \hline
        2006 & Wind & Design changes & 1999-2006 & Ops. & N & N & Y & N & \cite{Fingersh2006} \\
    \hline
        2012 & LED & Cost reductions & 2012 & Steps & N & N & Y & Y & \cite{ledcomv2} \\
    \hline
        2020 & LED & Impact of breakthroughs & 2002-now & Steps & (Y) & (Y) & Y & Y & $^\text{this}_\text{work}$\\
    \hline
\end{NiceTabularX}
\vspace{3mm}
\caption*{The \textit{Base} column describes the base unit for computations. This can either be manufacturing steps ("Steps"), cost reductions ("Red.") or total cost of operation ("Ops."). The \textit{Parameters} column describes the parameters considered. Market parameters include market size, manufacturing equipment discount, economies of scale, etc. Economy parameters include taxes, subsidies, debt, inflation, etc. "Global" parameters include resource cost, labour cost, overall yield, etc. "Step" parameters include equipment performance, equipment cost, step yield etc. Models self-described as \textit{"bottom-up"} consider parameters $G,S$. Note that those models only capture the manufacturing situation a certain point in time. The exception to this are the model \cite{Elia2020} and the present model. Abbreviations: Solar Photovoltaics, excluding perovskite cells (Solar), integrated circuits (IC), base level for calculations (Base), Parameters: market (M), economy (E), "global" (G), "step" (S). Parameters not considered directly but instead through their impact on other parameters denoted (Y).}
\label{tab:costmodels}
\end{table}

While generally comprehensive, the model is limited by the availability of performance data for the most recent manufacturing equipment, as well as the confidential nature of cost structures of Asia-based manufacturing lines. A complete treatment of the limitations of the model is provided in \cref{sec:costmodel_limitations}.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.85\textwidth]{./figures/SEMI_pyramid.pdf}
    \caption{\textit{SEMI Standard 35} pyramid showing variables that enter into the computation of manufacturing cost. Variables not captured by our model are greyed out. Source: Adapted from \cite{standardE35} and \cite{helin200618th}.}
    \label{fig:semi_pyramid}
    %\vspace{-10pt}
\end{figure}

\clearpage
\section{Methods: Cost Model Limitations}
\label{sec:costmodel_limitations}

\cref{sec:costmodel_structure} provided additional information on the structure of the cost model. As illustrated in \cref{tab:costmodels}, our model is comprehensive in its coverage of market, economic, manufacturing step and global manufacturing parameters. However, the model has limitations which must be considered when interpreting the results.

\textbf{Availability of Equipment Performance Data}

The data for the inputs used in the model were obtained through a combination of scientific literature, conference proceedings, industry reports, industry periodicals, manufacturing equipment datasheets and experts interviews. The model was therefore set up to consider four light-emitting diode chip architectures: A classical light-emitting diode chip architecture with lateral current spreading, two diﬀerent high-power thin-ﬁlm ﬂip-chip architectures and a chip-scale package thin-ﬁlm ﬂip chip architecture, all shown in \cref{sec:manufacturing_steps}.

Unfortunately, despite the multitude of data sources, important performance parameters of  manufacturing equipment relevant to the latest generation of chips, such as those shown in \cref{fig:manuf_vtf_2012-1}-\cref{fig:manuf_csp_2020-2}, were unavailable due to their highly proprietary nature. Since full coverage of manufacturing equipment performance parameters could therefore not be assured for all chip architectures, costs were computed only for the aforementioned classical chip architecture.

\textbf{Lack of Data for Calibration}

While the model is setup up to capture the cost structure of a \textit{"virtual"} North American factory, the majority of production capacity is presently located in Asia. There, the share of domestically produced manufacturing equipment has increased from 10\% in 2008 to 65\% in 2015 \cite{tseng2015fab}. Recent investments into factories for high-power light-emitting diodes include Osram's Kulim plant in Malaysia \cite{osram2020kulim} and Lumileds' Penang plants in Malaysia \cite{lumileds2020penang}. In the production of low-power light-emitting diodes, the world's largest producer of LED packages by volume, MLS Co. operates plants in mainland China \cite{mls2020dev}.

This means that it is difficult to validate model outputs directly with the help of industry data, which is provided only for Asia-based manufacturing lines. We compare cost model data to a selection of industry data in \cref{sec:cost_model_calibration} and provide an interpretation of the differing cost structure. Generally, we found that our model provides the best-case scenario for LED manufacturing, where the most performant (highest throughput, highest yield) equipment is used in each year. Interviews revealed that this is rarely the case in practice. One reason for this is economic pressure on western manufacturers after the introduction of significant Chinese subsidies as part of the country's Twelfth Five-Year Plan \cite{sanderson2014light}.

However, because the purpose of our model is to show the impact of changes in the design and manufacturing process of devices, the geographic location and associated potentially lower overhead costs or slower adoption of specific equipment are of secondary importance to our model. This is why a US based manufacturing environment was assumed.

\section{Manufacturing Process Steps by Chip Architecture}
\label{sec:manufacturing_steps}

@@

@@

\newpage
\section{Cost Model: Sensitivity Analysis}

\begin{figure}[ht!]
	\centering
    \includegraphics[width=\textwidth]{./figures/costmodel_sensitivity.pdf}
	\caption{\textbf{Sensitivity analysis for selected cost model parameters.} Panel a represents the tested range of variation of parameters indicated on horizontal axis along with their respective measurement units, as represented in Supplementary Table 3. Panel b shows the resulting range of variation in the manufacturing cost. Ranges are represented for three years considered in the cost model, where corresponding data is available, by whiskers of the following colour: 2003 – blue, 2012 – green, 2020 – red. Note the overall trend of decreasing sensitivity to cost model parameters as the number of die per wafer increases over time. Abbreviations: resp. - respective; Yield Sav. from Insp. - yield savings from inspection.}
	\label{fig:sensitivity}
\end{figure}

@@

\begin{table}[]
\centering
\small
    \caption{\textbf{List of cost model parameters considered in sensitivity analysis.}}
    \begin{NiceTabularX}{1.1\textwidth}{ |X|X|l|l|l|l|l|l|X|}
        \hline
            \textit{Parameter} & \textit{Unit} & \textit{2003} & $\pm [\%]$ & \textit{2012} & $\pm [\%]$ & \textit{2020} & $\pm [\%]$ & Source \\
        \hline
            Cleanroom Cost & \text{USD}/m$^2$ & 3000 & +16,-16 & 3000 & +16,-16 & 3000 & +16,-16 & \cite{mddi1997cleanroom}\cite{ledcomv2} \newline \cite{bakshi2009euv}\cite{gajera2006process} \\
        \hline
            Labor Rate & \$/h & 30 & +50,-50 & 30 & +50,-50 & 30 & +50,-50 & I \\
        \hline
            Equip. Discount & \% of \text{USD} & 0\% & +25,-25 & 0\% & +25,-25 & 0\% & +25,-25 & I \newline \cite{Appleyard_2001} \\
        \hline
            Overall Yield & \% & 31\% & +25,-25 & 57\% & +25,-15 & 74\% & +25,-25 & I, \cite{lumi2012yield}\cite{ledsmag2012} \newline \cite{systemplus2015reverse}\cite{ledcomv2} \\
        \hline
            Inspec. Yield Savings & \%/inspec. & 0.5\% & +80,-80 & 0.5\% & +80,-80 & 0.5\% & +80,-80 & \cite{mckinseyyield} \\
        \hline
            Overall Throughput & UPH or h$^{-1}$ & varies & +50,-50 & varies & +50,-50 & varies & +50,-50 & Equipment Datasheets \\
        \hline
            Wafer Diameter & mm & 100 & +0,-50 & 150 & +33.3,-33.3 & 200 & +0,-25 & I and \cref{fig:wafers} \\
        \hline
            Edge+Street Width & mm & 7 & +0,-50 & 5 & +40,-0 & 5 & +40,-20 & \cite{ledsmagexclusion}\cite{rubiconexclusion} \newline \cite{xiamenexclusion}\cite{american2007annual} \\
        \hline
            Cutting Width & $\mu$m & 100 & +50,-25 & 75 & +33.3,-20 & 20 & +300,-50 & \cite{masaki2000division}\cite{ils2005width} \newline \cite{photonics2010width}\cite{discowidth} \\
        \hline
            Metal Prices & \text{USD}/kg & varies & +50,-0 & varies & +50,-0 & varies & +50,-0 & Datasheets \\
        \hline
            Electricity Price & \text{USD}/kWh & 0.07 & +50,-50 & 0.08 & +50,-50 & 0.06 & +50,-50 & \cite{eia2000electric}\cite{eia2019electric} \\
        \hline
            Saph. Subst. Price & \text{USD} & 40 & +12.5,-12.5 & 10 & +100,-20 & 3 & +66.6,-33.3 & \cref{fig:wafers} \\
        \hline
            Phosphor Prices & \text{USD}/g & varies & +50,-50 & varies & +0,-0 & varies & +0,-0 & I, \cite{yole_phosphor_2012}\cite{yole2017phosphor} \\
        \hline
        \end{NiceTabularX}
    \vspace{2mm}
    \caption*{The results of the sensitivity analysis for the parameters in this table are presented in \cref{fig:sensitivity}. Note: Baseline values are provided only for \textit{global} model parameters in the columns \textit{2003, 2012, 2020}. For \textit{non-global} model parameters (those which vary by equipment) only the variation, in percent of the individual baseline values is provided. The corresponding units for the values in columns \textit{2003}-\textit{2020} are indicated in the column \textit{Units}. The variation employed during the sensitivity analysis is provided in the columns $\pm [\%]$. Abbreviations: UPH - units per hour; Equip. Discount - equipment discount (sales rebate for large purchases); Inspec. Yield Savings - yield savings from inspection (early detection and alleviation of issues in the manufacturing workflow); Saph. Subst. - Sapphire Substrate. ’I’ in the Source column indicates expert interviews as the source of information.}
    \label{tab:sensitivity}
\end{table}

\clearpage
\section{Cost Model: Comparison with Industry Data/Projections}
\label{sec:cost_model_calibration}

In order to validate the outputs of the cost model regarding the evolution of the cost structure of low-to-mid-power classic chip GaN phosphor-converted white LED, we conducted a comparison with previously reported U.S. DOE calculations and projections based on the LEDCOM model and industry data provided as part of industry round table discussions. This comparison is shown in \cref{fig:costmodel_calibration}.

We note some differences between the results of our model and the cost structure reported or projected by the DOE. For instance, the share of the epitaxy step is consistently larger in the DOE data. This can in part be explained by our model relying on state-of-the-art equipment at a virtual U.S.-based manufacturing location, while industry might not run low-power and mid-power chip production on these, more expensive, reactors. In addition, the manufacturing lines of the majority of manufacturers are located in Asia. We also note that the share of the sapphire substrate price in the DOE data is much larger than in our model, which can in part be explained by the overestimation of the actual price of sapphire wafers in earlier projections made for 2015-2020. Finally, the relative importance of the packaging part of the manufacturing process is very similar in our model and in the DOE results in 2012. However, in the DOE projections it significantly decreases by 2020, while in our model it retains and even increases its share of the total cost. This trend has been independently confirmed by researchers and industry reports on wafer-level packaging \cite{Lee2011WPL,Xie2013,ledsmag2017WLP}, showing better performance of our cost model compared to earlier DOE model projections.

@@

\clearpage
\section{Full List of Sources for Figures}
\label{sec:sources}

\begin{table}[h!]
    \caption{List of sources for Figure 1 organized by technology.}
    \begin{NiceTabularX}{\textwidth}{|l|X|}
    \hline
    \textit{Technology} & \textit{References} \\
    \hline
    LED & Own research, compare \cite{zenodo_weinold_led_history} \\
    \hline
    CFL (<1984) & \cite{Bouwknegt1982}\cite{Vrenken1983} \\
    \hline
    CFL (1984-2011) & \cite{eger2018origin} \\
    \hline
    CFL (>2011) & \cite{Guan2015} \\
    \hline
    Fire, Incandescent, HID & \cite{azevedo2009transition} augmented by own calculations based on \cite{benesch1905beleuchtungswesen} \\
    \hline
    Max. efficacy & \cite{Murphy2012} \\
    \hline
    \end{NiceTabularX}
    \vspace{2mm}
    \caption*{Abbreviations: CFL - Compact Fluorescent Lamp, HID - High-Intensity Discharge}
\end{table}

@@

@@

@@

\printbibliography[title=Supplementary References]
